---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# surveymonkey

<!-- badges: start -->
<!-- badges: end -->

This package provides access from R to the SurveyMonkey API.  You can browse your surveys, pick one to fetch, and then, most importantly, parse the fetched JSON result into a tidy data.frame.

### Why this is useful

Compared to downloading .csv files manually:

- No fussing with the web browser or logging in
- Column names are handled appropriately - not split over two rows
- The columns are factors when appropriate, with the levels ordered based on the sequence the answers appear in the survey.
  - And they have knowledge of all choices that were offered on the survey, even if they're missing from the data because no one selected them.  These values would be absent from a .csv download.

**This package is still in development and may not work on 100% of use cases**, but will fetch the typical survey correctly and is worth a shot before you go to wrangle yet another .csv or .xlsx export.


### What's working

Most data is  pulled from the API: all responses are pulled, as well as most metadata like custom variables and response start/end timestamps.   All substantive question types should be currently implemented.

### What's missing
Some uncommon question types may not yet be accounted for.  For instance, no special effort was made to address image or upload question types.

If you have a use case for something that isn't currently pulled through, please open an issue describing your situation & question type.


## Installation

This package is not yet on CRAN.  Install from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("tntp/surveymonkey")
```
 
## Usage

### Authentication

*Have an opinion about OAuth procedures?  If you can improve this guidance, please open an issue with your suggestions.*

#### Get your OAuth token

You'll need an OAuth token, and for that you'll need to set up an app.

Log in to SurveyMonkey in your browser, then navigate to https://developer.surveymonkey.com/apps.  Create an app.  It should be private, and you should enable the relevant scopes like View Surveys, View Collectors, View Contacts, View Responses, View Response Details, etc. (if you get a 403 error when you try to browse surveys, try enabling other relevant scopes).  Deploy the app.

Now look at the settings page for your app and take note of the "Access Token" field, which should contain a very long character string.

#### Add your OAuth token to your .Rprofile

Add the SurveyMonkey account's OAuth token to your .Rprofile file.  To open and edit that file, run `usethis::edit_r_profile()`, then add a line like this:
`options(sm_oauth_token = "kmda332fkdlakfld8am7ml3dafka-dafknalkfmalkfad-THIS IS NOT THE REAL KEY THOUGH")`.

Except use the OAuth token listed on your app's settings page, obtained in the previous step.

Restart R for this change to take effect.

If this is all set up successfully, the token will print when you run `getOption("sm_oauth_token")`.  Guard this token: don't share it and don't commit it in any repository.

*Developer's note: That's how a user can get a single account's OAuth token.  It might be preferable if users could authenticate from within R.  If someone has guidance for how users should obtain their OAuth token more generally, please submit a pull request or comment in an issue.*


### Browsing your surveys

You'll need the ID number of the survey you wish to fetch.  Find it by browsing your surveys like this:
```{r, eval = FALSE}
surveys <- browse_surveys(200) # see your most recent 200 surveys
```
Then run `View(surveys)` and start typing the name of the survey you want into the search window at the top right of the RStudio data.frame viewer.  That will get you the survey's ID number.  Copy it.

### Fetching a survey

Get the survey data like this:
```{r, eval=FALSE }
a_survey_obj <- fetch_survey_obj(123456789) # your survey's ID goes here
```

This returns a list of data about the survey.  It's useful for developers to explore, but not so much for most users.  Keep going to the next step.

### Parsing the survey into a data.frame

This is the actual good part.
```{r, eval = FALSE}
survey_df <- parse_survey(a_survey_obj)
```

That will give you a tidy data.frame with all of your responses.

In the future you can run it all as one command:

```{r, eval = FALSE}
survey_df <- 123456789 %>%
  fetch_survey_obj %>%
  parse_survey
```

## API considerations
Your account will likely be limited to 500 hits per day to the API.  This package will print reminders of how many calls you have left in the day.  The main thing to keep an eye on is respondent counts; as only 100 responses can be fetched per API call, a survey with X respondents will make at least X/100 calls to the API.


# Contact

Something may not work, because this is an in-development package.  Or you may have an idea for something that hasn't been implemented yet.  In either case, please create an issue in the GitHub issue tracker!
